DATA 605 Week 15

### 2019-05-08 DATA605/MSilva_Assign15.Rmd
More work on Assigment 15

>>>>

Given the set of (X, Y) points we need to find the best fit line $\hat{Y}= a + bX$ such that the sum of square errors in Y ($\sum{(Y-\hat{Y})^2}$) is minimized.

$$SSE=\sum_{i=1}^{n}{(Y_i-\hat{Y})^2} = \sum_{i=1}^{n}{(Y_i-a - bX_i)^2}$$

SSE will be minimized when $\frac{d{SSE}}{da} = 0$ and $\frac{d{SSE}}{db} = 0$.

$$\frac{d{SSE}}{da} = \sum_{i=1}^{n}{-2(Y_i-a - bX_i)}=2\bigg(na+b\sum_{i=1}^{n}{X_i}-\sum_{i=1}^{n}{Y_i}\bigg)=0$$

Dividing by 2 and solving for a

$$a = \bar{Y}-b\bar{X}$$

$$\frac{d{SSE}}{db} = \sum_{i=1}^{n}{-2X_i(Y_i-a - bX_i)}=\sum_{i=1}^{n}{-2(X_iY_i-aX_i-bX_i^2)}=0$$

Substituting in a gives

$$\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y}-bX_i\bar{X}-bX_i^2)}=0 \rightarrow \sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}-b\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}=0$$

Solving for b gives

$$b=\frac{\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}}{\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}}$$

```{r}
X <- c(5.6, 6.3, 7, 7.7, 8.4)
Y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

X_bar <- mean(X)
Y_bar <- mean(Y)

b <- sum(X * Y - X * Y_bar) / sum(X ^ 2 - X * X_bar)

a <- Y_bar - b * X_bar
```

The least squares line is $\hat{Y}$ = `r a` + `r b` $X$

Double checking my work
```{r}
lm(Y ~ X)
```
